Geometric variational problems

### 2020-11-12 NSF2020/Description/specificaim2_ver2.tex

We will compare the behavior of our particle-based vesicles in Stokes flow to well-established results.
\todo{What are the known results?}
To test area incompressibility, we form the bilayer midplane by fitting to the particle centers. Under moderate shear rates,
the vesicle elongates and takes on an elliptical shape. We will study the increase in midplane area as a function of shear rate.
Additionally, we will test for tangential divergence of velocity.
Researchers have devised to model area incompressible and
volume conserving membranes
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
Our preliminary results show that the particle-based bilayers have the same area modulus, about 60 \kBT,  as that of real lipid bilayer membranes,
and so we expect realistic results from our flow simulations with regard to area incompressibility.  It is noteworthy that
area incompressibility is some sense baked into the model through the hydrophobic attraction.

@@

\begin{wrapfigure}[13]{r}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{figures/BIError.pdf}}
\caption{
\label{fig:bierror}  
The numerical error in integral representations blows up near boundaries.
}
\end{wrapfigure}

Finally, the particle-based vesicle bilayers have two distinct leaflets.
The inner and outer leaflets consists of the particles in contact with the vesicle interior and 
surrounding fluid respectively. Since there the separation between these layers is roughly fixed,
about 2 nm, certain fluid mechanical effects arise that are not present in zero-thickness membrane models. 
An important effects we observe is that the leaflets slide against one another under shear flow. 
This implies that part of the viscous dissipation in the aqueous phase is enhanced by intermonolayer friction
\cite{SHKULIPA2005823, ShkulipaThesis}. We will characterize the intermonolayer
slip as a function of shear rate, vesicle diameter and particle geometry.

\textbf{Novel reciprocal relations.}
We use a boundary integral representation for the velocity field. Due to singular behavior of the
kernels, the boundary integrals can have large numerical epresentation without the aid of specilized
quadrature routines, leads to large numerical errors when evaluating the the hydrophobic
stress \eqref{stress} along the particle boundaries (see Figure \cite{fig:bierror}).
We prove a, to our knowledge new, result concerning the evaluation of Maxwell-type stresses. 
\begin{proposition}
  \label{prop:recip}
  Suppose that $u = \sum_{i=1}^N u_i$ where $u_i$ is a solution of the screened Laplace equation
  in $\mathbb{R}^3 \setminus P_i$ for $i=1,\dots, N.$ Then 
  \begin{equation}
    \label{eq:reciprocal}
{\bf F}_i = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_i = \sum_{j \neq i} \int_{\partial P_i} {\bf x} \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, \quad i = 1\dots, N.
\end{equation}
where $\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump in stress across $\partial \Omega.$ 
\end{proposition}

The identity in Proposition \ref{prop:recip} gets around this critical issue by utilizing the fact that $u_j$, $j\neq i$ is smooth and the normal derivative
of  $u_i$ continuous across $\partial P_i$. Despite the large gradient errors in Figure \ref{fig:bierror}, the force and torque
calculated by \eqref{eq:reciprocal} achieved $10^{-4}$ accuracy.

>>>>

We will compare the behavior of our particle-based vesicles in Stokes flow to well-established results.
\todo{What are the known results?}
Researchers have devised models to enforce area incompressibility and
volume conservation in vesicle membrane dynamics. 
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
Our preliminary results show that the particle-based bilayers have the same, large area modulus as that of real lipid bilayer membranes,
and so we expect realistic results from our flow simulations.
Under moderate shear rates, the vesicle elongates and takes on the tank-treading, elliptical shape.
We can directly check for area compressibility conservation properties from the particle simulation,
by tracking the change in area density of the midplane as a function of shear flow rate,  
along with evaluating the tangential divergence of the velocity.

@@

\begin{wrapfigure}[16]{r}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
\caption{
\label{fig:bierror}  
The false color map shows how numerical quadrature of layer
potentials loses accuracy near the particle boundaries.
The color bar is for $\log_{10}.$ 
}
\end{wrapfigure}

Finally, the particle-based vesicle bilayers have two distinct leaflets,
consisting of the particles in contact with the vesicle interior and 
surrounding fluid respectively. Since there is a nonzero separation between these layers,
one important effect we observe is that the leaflets slide against one another under shear flow. 
This implies that part of the viscous dissipation in the aqueous phase is enhanced by intermonolayer friction
\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity jump boundary condition
and leads to multiple solution states \cite{schwalbe_vlahovska_miksis_2010}.
In the present proposal, intermonolayer slip is an artifact of monolayer independence  
and we can control for slip as a function of shear rate, vesicle diameter and particle geometry.

\textbf{Novel reciprocal identities.}
In the course of our study, we have found it useful to devise certain reciprocal identities that
allow us to calculate forces accurately without having to resort to overly complicated numerical integration schemes. 
At issue is that there is a break-down in numerical accuracy for the gradient of water activity near the particle
boundaries (see Figure \ref{fig:bierror}).
This is unfortunate the boundary is exactly where we need the gradients to be accurate in order for the formulas \eqref{forceandtorque}
to be meaningful. To get around this issue, we prove the following identity
  \begin{equation}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S
  \end{equation}
for $i = 1\dots, N$. Here $u = \sum_{i=1}^N u_i$ is a sum of solutions $u_i$ the screened Laplace equation,
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump across $\partial \Omega.$  It turns out that because of \cite{eq:reciprocal},
we no longer require knowledge of the gradient on the boundary, which is enormously beneficial for calculating force and torque.
